---
title: "Project_1_New"
author: "Sanjeev's SupeRstars: Aaron Zelmanov, Muhammad Hafizudeen Mohamad Saman, Nakul Chadha, Kendall Cohen, Michael Geraci, Michael Zhang"
date: "2/15/2021"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setup 
hotel <- read.csv("hotel_bookings.csv")
library("ggplot2")
library("dplyr")
library("lubridate")
library("zoo")
library("priceR")

#factorizing 
hotel$is_canceled <- as.factor(hotel$is_canceled)
hotel$hotel <- as.factor(hotel$hotel)
hotel$arrival_date_year <- as.factor(hotel$arrival_date_year)
hotel$arrival_date_month <- as.factor(hotel$arrival_date_month)
hotel$arrival_date_week_number <- as.factor(hotel$arrival_date_week_number)
hotel$is_repeated_guest <- as.factor(hotel$is_repeated_guest)
hotel$reservation_status <- as.factor(hotel$reservation_status)
hotel$customer_type <- as.factor(hotel$customer_type)
hotel$deposit_type <- as.factor(hotel$deposit_type)
hotel$reserved_room_type <- as.factor(hotel$reserved_room_type)
hotel$assigned_room_type <- as.factor(hotel$assigned_room_type)
hotel$distribution_channel <- as.factor(hotel$distribution_channel)
hotel$market_segment <- as.factor(hotel$market_segment)
hotel$country <- as.factor(hotel$country)
hotel$meal <- as.factor(hotel$meal)
hotel$company <- as.factor(hotel$company)
hotel$agent <- as.integer(hotel$agent)
hotel$has_agent <- ifelse(is.na(hotel$agent) == TRUE, 0, 1)
hotel$arrival_date_day_of_month <- as.factor(hotel$arrival_date_day_of_month)
```
## {.tabset}
### Introduction 

**Context:** We are working with the hotel_booking.csv dataset to better understand hotel bookings and cancellations. This dataset includes information about: dates, demographics, customer types, reservation details, cancellations, etc.

**Audience:** Hotel managers who are worried about hotel cancellations

**Key Question:** What is the root cause of cancellations and how can we mitigate this issue? 

### Are Cancellations a Problem? 
In order to understand if cancellations are a real problem, we must explore our data! We will analyze the number of cancellations and proportion of cancellations over time and then figure out how much this is affecting our revenue 

```{r}
cancellationNumberPlot <- ggplot(data=hotel, aes(x=is_canceled, fill=is_canceled)) + geom_bar() + xlab("Not Cancelled or Cancelled") + ylab("Count") + ggtitle("Count of Reservations Not Cancelled vs Cancelled") + labs(caption=("Number of Cancellations = 44,224. Total reservations = 119,390")) + theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5))
cancellationNumberPlot + guides(fill = FALSE)
```

```{r}
cancellationPropPlot <- ggplot(data=hotel, aes(x=is_canceled, y=(..count..)/sum(..count..), fill=is_canceled)) + geom_bar() + xlab("Not Cancelled or Cancelled") + ylab("Proportion") + ggtitle("Proportion of Reservations Not Cancelled vs Cancelled") + labs(caption=("More than a third of bookings (37%) were cancelled, which is a significant amount")) + theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5)) + guides(fill = FALSE)
cancellationPropPlot
```

```{r, warning = FALSE, warn.conflicts = FALSE, dplyr.summarise.inform = FALSE}
cancellations2015Table <- hotel[hotel$is_canceled==1 & hotel$arrival_date_year==2015,]
cancellations2015 <- nrow(cancellations2015Table)
cancellations2016Table <- hotel[hotel$is_canceled==1 & hotel$arrival_date_year==2016,]
cancellations2016 <- nrow(cancellations2016Table)
cancellations2017Table <- hotel[hotel$is_canceled==1 & hotel$arrival_date_year==2017,]
cancellations2017 <- nrow(cancellations2017Table)

bookings2015Table <- hotel[hotel$arrival_date_year==2015,]
bookings2015 <- nrow(bookings2015Table)
bookings2016Table <- hotel[hotel$arrival_date_year==2016,]
bookings2016 <- nrow(bookings2016Table)
bookings2017Table <- hotel[hotel$arrival_date_year==2017,]
bookings2017 <- nrow(bookings2017Table)

table1 <- data_frame(year = c(2015,2016,2017), count = c(cancellations2015,cancellations2016,cancellations2017), prop = c(cancellations2015/bookings2015,cancellations2016/bookings2016,cancellations2017/bookings2017))

cancellationNumberByYear<- ggplot(data=table1, aes(x=year,y=count,fill=year)) + geom_bar(stat='identity') + ggtitle("Number of Cancellations by Year") + labs(x="Year", y="Count",caption=("The number of cancellations increased from 2015 to 2016 but has decreased since then")) + theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5)) + guides(fill = FALSE)
cancellationNumberByYear

hotel$is_canceled_2 <- as.numeric(hotel$is_canceled)
hotel$month_year <- as.yearmon(paste(hotel$arrival_date_month, hotel$arrival_date_year))
hotel %>% group_by(month_year) %>%
   summarize(sum_canceled=sum(is_canceled_2)) %>% ggplot() + geom_point(aes(x=month_year, y=sum_canceled)) + geom_line(aes(x=month_year, y=sum_canceled)) + ggtitle("Number of Cancellations by Month over Time") + labs(x="Month and Year", y="Count",caption=("The number of cancellations are variable (seasonality + idiosyncratic effects) but show a somewhat increasing trend")) + theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5))

cancellationPropByYear <- ggplot(data=table1, aes(x=year,y=prop, fill=year)) + geom_bar(stat='identity') + ggtitle("Proportion of Cancellations by Year") + labs(x="Year", y="Proportion",caption=("The proportion of cancellations decreased from 2015 to 2016 but has slightly increased since then")) + theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5)) + guides(fill = FALSE)
cancellationPropByYear

hotel %>% group_by(month_year) %>%
    summarize (n=n()) %>% mutate(prop_canceled=sum(n)/n) %>% ggplot() + geom_point(aes(x=month_year, y=prop_canceled)) + geom_line(aes(x=month_year, y=prop_canceled)) + ggtitle("Proportion of Cancellations by Month over Time") + labs(x="Month and Year", y="Proportion",caption=("The proportion of cancellations are also variable but show a somewhat decreasing trend")) + theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5)) 
```

*Based on the findings, we know that from 2015-2016, the number of bookings was outpacing the number of cancellations because the proportion was decreasing. Since then, the number of cancellations (although less than before) has been outpacing the number of bookings*



**Are we losing money because of the cancellations?**

```{r}
table_mean_adr_canceled <- tapply(hotel$adr, hotel$is_canceled, mean, na.rm = TRUE)
df <- as.data.frame(table_mean_adr_canceled)
colnames(df) <- c("Mean ADR")
df$`Mean ADR` <- as.numeric(as.character(df$`Mean ADR`))
df$`Mean ADR` <- round(df$`Mean ADR`, 2)
df
```

*The table above shows the mean ADR for if the reservation was not cancelled and cancelled. As we can discern, the mean ADR is 5 dollars higher for cancelled reservations. Although this may not constitute a large difference, if extrapolated to the entire data set of almost 120,000 reservations, a large amount of money is lost. Namely, if we multiply the mean ADR by the total number of cancellations:*

```{r}
mean_adr_per_year <- aggregate(adr ~ arrival_date_year + is_canceled, data = hotel, mean)
graph <- ggplot(data = as.data.frame(mean_adr_per_year), aes(x=arrival_date_year, y = adr, color = is_canceled)) + geom_point(size = 6) + ggtitle("Mean ADR per Arrival Year") + labs(x = "Arrival Year", y="Mean ADR", caption=("As the year increases, the mean ADR for cancellations increases.")) + theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5))
mean_adr_per_year
graph
```


```{r}
priceR::format_dollars(104.96 * 44224)
```

*As demonstrated, more than $4.5 million dollars is lost.*

```{r}
table_deposit_type <- table(hotel$deposit_type, hotel$is_canceled)
addmargins(table_deposit_type)
```

```{r}
prop_table <- prop.table(table_deposit_type)
addmargins(prop_table)
```
*We can make several observations from the two tables above. First, cancellations make up 37% of all reservations. However, of those cancelling, there is a tremendous difference in non-refundable reservations. Although customers pay a non-refundable deposit, this does not constitute the full reservation amount, meaning that the hotel will still lose the money.*

### Causes of the Cancellation Problem 
Now that we know cancellations are indeed a problem for hotel managers, we must understand what is causing them! We will begin my understanding the correlations of various variables with cancellations and then understand how these variables relate to a cancelled booking. Thereafter, we will explain our results in order to better illustrate what we are seeeing in the data. 
```{r, warning = FALSE}
hotel$arrival_date_year_2 <- as.numeric(hotel$arrival_date_year)
hotel$arrival_date_month_2 <- as.numeric(hotel$arrival_date_month)
hotel$arrival_date_week_number_2 <- as.numeric(hotel$arrival_date_week_number)
hotel$is_repeated_guest_2 <- as.numeric(hotel$is_repeated_guest)
hotel$arrival_date_day_of_month_2 <- as.numeric(hotel$arrival_date_day_of_month)

correlations <- as.data.frame(cor(hotel[sapply(hotel, function(x) is.numeric(x))])[16,])
correlations %>% ggplot(aes(x=reorder(rownames(correlations), correlations[,1]), y=correlations[,1])) + 
  geom_bar(stat = 'identity') + ggtitle("Correlations with Cancellations") + labs(x="Variables", y="Cancellations",caption=("The bar graph above displays the hotel_bookings variables that have a correlation with cancellations in increasing order")) + theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90))


# (Kendall + Mike + Michael) plotting the highly correlated variables with cancellations and explaining the correlations. Then, buildiing additional graphs to further analyze
```

**Previous Cancellations Indicators**

Proportion of those with previous cancellations that have cancelled
``` {r}
has_canceled <- hotel[hotel$previous_cancellations > 0,]
prop.table(table(has_canceled$is_canceled))
ggplot(has_canceled, aes(is_canceled, fill= is_canceled))+ geom_bar() + labs(x="Canceled Reservation", y= "Count of Canceled Reservations") + ggtitle("Reservations canceled by those who have canceled in the past")

```

*Based on this finding we see that if someone has canceled their reservation in the past, they are highly likely to cancel once again. This is an example of the motif: "The best indicator of future behavior is past behavior*

Proportion of those with special requests who have canceled
```{r}
special_requests <- hotel[hotel$total_of_special_requests > 0,]
prop.table(table(special_requests$is_canceled))
ggplot(special_requests, aes(is_canceled, fill= is_canceled))+ geom_bar()
```


```{r}
# Lead Time 
summary(hotel$lead_time)
hotel$lead_time_range <- cut(hotel$lead_time, 
                             breaks = c(0, 18, 69, 160, 738), right = FALSE, 
                             labels = c( "Lowest", "Low/Middle", "High/Middle", "Highest"))

plot(hotel$lead_time_range)

ggplot(hotel, aes(x=lead_time_range, fill=is_canceled)) + geom_bar()

ggplot(hotel, aes(x=is_canceled, fill=lead_time_range)) + geom_bar()

# LT Prop Table 1: Prop table showing percentage of each lead time group that canceled/ did not cancel. 1 indicates cancellation. 
prop.table(table(hotel$lead_time_range, hotel$is_canceled),1)

# Pie chart showing lead time to cancellations 
hotel_is_canceled <- hotel[!(hotel$is_canceled == 0),]
hotel_is_canceled %>% group_by(lead_time_range) %>% summarize(n=n()) %>% ggplot(aes("", n, fill = lead_time_range)) + geom_bar(stat = 'identity', width = 1, color = "white") + coord_polar("y", start = 0) + theme_void() + labs(title = "Lead Time When Guest Cancelled")

# Pie chart showing lead time when NOT canceled
hotel_is_not_canceled <- hotel[!(hotel$is_canceled == 1),]
hotel_is_not_canceled %>% group_by(lead_time_range) %>% summarize(n=n()) %>% ggplot(aes("", n, fill = lead_time_range)) + geom_bar(stat = 'identity', width = 1, color = "white") + coord_polar("y", start = 0) + theme_void() + labs(title = "Lead Time When Guest Did NOT Cancel")

ggplot(hotel, aes(lead_time,  colour = is_canceled)) + geom_bar()

ggplot(hotel_is_not_canceled, aes(arrival_date_week_number, lead_time,  colour = is_canceled)) + geom_point()

# There is greater lead time in the middle of the year (summer)


```
LT Prop Table 1 and the above pie charts show that there is a very obvious relationship between lead time and cancellation. The further in advanced a person books, the more likely that they canceled. Of all the people who booked in the "Highest Lead Time" bucket, 55.3% ended up canceling. This is vastly different from people in the "Lowest Lead Time" bucket, where only 14% canceled. We also separated people who booked on the same day, and found that 6.8% of people canceled. This was surprising to the team because although 6.8% is relatively low, we did not expect people to change their minds so quickly.

**Presence of Agent**

*Now, we want to look if having an agent have an affect on the rate of cancellation*
```{r}
## plot of people who does not have agent
ggplot(data=hotel[hotel$has_agent == 0,], aes(x=is_canceled,fill= as.numeric(is_canceled))) + ggtitle("People who Does not have Agent") + xlab("Cancelled Reservation") + geom_bar()

##proportion of people who canceled in the group of people who doesnt have agent
sum(hotel$has_agent == 0 & hotel$is_canceled == 1)/sum(hotel$has_agent == 0)


## plot of people who have agent
ggplot(data=hotel[hotel$has_agent == 1,], aes(x=is_canceled,fill= as.numeric(is_canceled))) + ggtitle("People who have Agent") + xlab("Cancelled Reservation") + geom_bar()

##proportion of people who canceled in the group of people who have agent
sum(hotel$has_agent == 1 & hotel$is_canceled == 1)/sum(hotel$has_agent == 1)
```
*From the graph above, more than 39% of people who have an agent canceled their reservation compared to only 24.7% in the people who does not have an agent. This suggests that people who have an agent have more tendency to cancel their reservations compared to the ones who does not have any agent.*

### Solutions to the Cancellation Problem

**Solution for Previous Cancellations Problem**

The main reason that cancellations are a problem is because if the hotel is at capacity, or if the most profitable rooms have been booked, then you are missing out on lost revenue from the customers that would have booked and not cancelled. One solution to this problem would be to force those that have cancelled before to put down a deposit when they reserve in order to mitigate any losses that may occur from the lost revenue of other potential customers. Another solution is to reserve rooms for customers that have not cancelled in the past. You would apply Newsvendor Problem solving logic to determine the number of rooms that you leave reserved. 

**Analysis of Special Requests Indicators**

From our analysis, we view that those who have contacted the hotel for special requests are much less likely to cancel their reservation than a client who has not made any special requests. The hotel should take note of this whenever someone makes a request. However, correlation is not causation and it is not evident whether people decide they are not canceling before they make special requests or if they do not cancel because they have made those requests.

**Lead Time Analysis**

In our lead time analysis, there was a clear relationship between how far in advance a person booked and whether they cancelled their reservation or not. The trend showed that the further in advance that someone booked, the more likely they were to cancel their reservation. This is a difficult problem to have because for hotels reservations are beneficial for planning and gives the hotels a safeguard against the variance in room demand within any given week. Some possible solutions include, like the previous cancellations problem, to mandate that guests put down a deposit when they reserve a room far in advance, especially in our 
```{r}
# To be done later (after Wednesday)
```

