---
title: "Project 1"
author: "Aaron Zelmanov, Muhammad Hafizudeen Mohamad Saman, Nakul Chadha, Kendall Cohen, Michael Geraci, Michael Zhang"
date: "2/12/2021"
output:
  html_document:
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
# Data Reading
hotel <- read.csv("hotel_bookings.csv")
# Data Cleaning
hotel$hotel <- as.factor(hotel$hotel)
hotel$is_canceled <- as.logical(hotel$is_canceled)
hotel$arrival_date_year <-as.factor(hotel$arrival_date_year)
hotel$arrival_date_month <-as.factor(hotel$arrival_date_month)
hotel$meal <-as.factor(hotel$meal)
hotel$market_segment <-as.factor(hotel$market_segment)
hotel$country <-as.factor(hotel$country)
hotel$distribution_channel <-as.factor(hotel$distribution_channel)
hotel$is_repeated_guest <-as.logical(hotel$is_repeated_guest)
hotel$customer_type <-as.factor(hotel$customer_type ) 
hotel$reservation_status <-as.factor(hotel$reservation_status)
hotel$reserved_room_type <- as.factor(hotel$reserved_room_type)
hotel$assigned_room_type <- as.factor(hotel$assigned_room_type)
hotel$deposit_type <- as.factor(hotel$deposit_type)
```
## {.tabset}
### Basic Exploration

**Average number of people that stay at each hotel**
```{r}
hotel$num_people <- hotel$adults + hotel$children + hotel$babies
tableHotelInfo <- aggregate(num_people~hotel, data = hotel, mean)
tableHotelInfo$total <- c(sum(hotel$hotel == "City Hotel"), sum(hotel$hotel == "Resort Hotel"))
tableHotelInfo
```
*The table above shows the mean number of guests per reservation and total guests in each hotel from the data. We can imply through this data that the Resort Hotel is more family-friendly as the average reservation size is higher. Also, businesspeople tend to reserve the City Hotel as demonstrated by the less reservation size** 

**Hotel cancellations**
```{r}
plot(as.factor(hotel$is_canceled), xlab= "is_cancelled", ylab = "number of people")
nrow(hotel[hotel$is_canceled == 1,])/nrow(hotel)
```
*Table above shows the number of hotels cancelled and not cancelled. We can see that more than a third canceled their bookings. Total bookings = 119,390*

**Average lead time per hotel**

```{r}
avg_lead_time_table <- tapply(hotel$lead_time, hotel$hotel, mean, na.rm = TRUE)
avg_lead_time_table
```
*The table above shows the average lead time per hotel. We can see that people, on average, reserve the City Hotel almost 20 days prior to the Resort Hotel.*

**Number of people by market segment**

```{r}
#summary(hotel$market_segment)
#plot(hotel$market_segment)
market_seg_table <- table(hotel$market_segment, hotel$hotel)
addmargins(market_seg_table)
```
 
*The table above shows the number of people per hotel by market segment.*


**What times of the year are busiest?** 
```{r}
hotel$seasons <- ifelse(hotel$arrival_date_month == "December" | hotel$arrival_date_month == "January" | hotel$arrival_date_month =="February", "Winter", ifelse(hotel$arrival_date_month == "March" | hotel$arrival_date_month == "April" | hotel$arrival_date_month ==  "May",  "Spring", "Summer"))

library("ggplot2")
ggplot(hotel, aes(seasons)) + geom_bar()
```

*Chart above shows how reservations vary with seasons* 

**How about the arrival month? Are there any famous months for hotel reservation?**
```{r}
plot(hotel$arrival_date_month, las = 2)
table(hotel$arrival_date_month)
```
*The most famous moth for hotel arrival date is August. In the trend, we can see that the months with the lowest number of arrival months are months at the end and the beginning of the year, particularly from November-February*

### Cancellation Analysis 

**Cancellations by month**
```{r}
cancellationsbymonth <- tapply(hotel$is_canceled, hotel$arrival_date_month, mean, na.rm = TRUE)
barplot(sort(cancellationsbymonth), las = 2)
hotel %>% group_by(arrival_date_month) %>% summarise(num_cancel = sum(is_canceled)) %>% arrange(desc(num_cancel)) %>% head(10) %>% print %>% ggplot(aes(x = arrival_date_month, y = num_cancel, fill= num_cancel))+ geom_col()
```
*Most of the months have about 30-40% rate of cancellation, so we cannot definitely say for sure that month affect the rate of cancellation*

**Cancellations by hotel type**
```{r}
tableCancellationsbyType <- aggregate(is_canceled ~ hotel, data = hotel, sum)
cityHotels <- sum(hotel$hotel == "City Hotel", na.rm = TRUE)
resortHotels <- sum(hotel$hotel == "Resort Hotel", na.rm = TRUE)
tableCancellationsbyType$prop_cancelled <- c(round(33102/cityHotels, 2), round(11122/resortHotels, 2))
tableCancellationsbyType
```
*Table above shows the number of cancellations per hotel type and the proportion of each hotel's bookings that are cancelled* 

**Cancellations by market segment**
```{r}
tableCancellationsbyMarket <- aggregate(is_canceled ~ market_segment, data = hotel, length)
tableCancellationsbyMarket
```
*Table above shows the cancellations per market segment* 

**Cancellation by country**
```{r}
hotel %>% group_by(country) %>% summarise(num_cancel = sum(is_canceled)) %>% arrange(desc(num_cancel)) %>% head(10) %>% print %>% ggplot(aes(x = country, y = num_cancel, fill= num_cancel))+ geom_col()
```


